Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images

### 2017-01-16 paper.tex
minor updates in paper

% \tableofcontents
% \clearpage
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
  A  large  scale  feature  selection wrapper  is  discussed  for  the
  classification  of high  dimensional  remote  sensing. An  efficient
  implementation is proposed based on intrinsic properties of Gaussian
  mixtures models and  block matrix.  The criterion  function is split
  into two parts  : one that is  updated to test each  feature and one
  that needs to be updated only once per feature selection. This split
  saved  a  lot  of  computation  for each  test.   The  algorithm  is
  implemented in  \texttt{C++} and integrated into  the Orfeo Toolbox.
  It has been compared to  other classification algorithms on two high
  dimension remote sensing images. The  results show that the approach
  provides good classification accuracies with low computation time.
\end{abstract}

@@

The volume of the data is  increasing dramatically with respect to the
number of measurement  per pixel. The data volume  of an hyperspectral
image  is  typically several  hundreds  of  Gigabytes per  acquisition
($\approx$  300km$^2$). Multitemporal  data are  now available  freely
from  internet  stream  (see  for instance  the  Copernicus  data  hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of parallel approaches (multi-core, GPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  Space
Agency (CNES)~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  conveniently proposed  to  users/developpers through  several ``\emph{ready to use}'' modules.

@@

The resulting code is available as a remote module of the Orfeo toolbox on Github and makes it possible to process large high dimensional images efficiently. The \texttt{C++} code is freely available for download: \url{https://www.orfeo-toolbox.org/external-projects/}.

@@

    % \begin{align*}
    %     \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
    %     \log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
    %             \mathbf{u}_c^t          & \sigma^{(k)}_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %     \right|\right)
    %     +\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \alpha_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
    % \end{align*}
    % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediatly:
  \begin{align*}
    \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\sigma^{(k)}_c  - \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}  \mathbf{u}_c\right)\\
    &= \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\alpha_c\right)
  \end{align*}

>>>>

% \tableofcontents
% \clearpage
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
  A  large  scale  feature  selection wrapper  is  discussed  for  the
  classification  of high  dimensional  remote  sensing. An  efficient
  implementation is proposed based on intrinsic properties of Gaussian
  mixtures models and  block matrix.  The criterion  function is split
  into two parts  : one that is  updated to test each  feature and one
  that needs to be updated only once per feature selection. This split
  saved  a  lot  of  computation  for each  test.   The  algorithm  is
  implemented in  \texttt{C++} and integrated into  the Orfeo Toolbox.
  It has been compared to  other classification algorithms on two high
  dimension remote sensing images. Results show that the approach
  provides good classification accuracies with low computation time.
\end{abstract}

@@

The volume of the data is  increasing dramatically with respect to the
number of measurement  per pixel. The data volume  of an hyperspectral
image  is  typically several  hundreds  of  Gigabytes per  acquisition
($\approx$  300km$^2$). Multitemporal  data are  now available  freely
from  internet  stream  (see  for instance  the  Copernicus  data  hub
\url{https://cophub.copernicus.eu/}).  This very  large volume of data
requires   specific   computing  infrastructure.    High   performance
computing   is   actually   investigated   by   the   remote   sensing
community~\cite{christophe2011remote,plaza2011high}.  Main  issues are
related to the use of parallel approaches (multi-core, GPU, clusters) to
improve the  processing time, and  to the use of  streaming techniques
when data  does not fit in  memory.  One popular open  source software
solution  is  the Orfeo  Toolbox,  developed  by the  French  Space
Agency (CNES)~\cite{christophe2008orfeo}.   Streaming and  parallel computing
are  conveniently proposed  to  users/developers through  several ``\emph{ready to use}'' modules.

@@

The resulting code is available as a remote module of the Orfeo ToolBox on GitHub and makes it possible to process large high dimensional images efficiently. The \texttt{C++} code is freely available for download: \url{https://www.orfeo-toolbox.org/external-projects/}.

@@

    % \begin{align*}
    %     \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
    %     \log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
    %             \mathbf{u}_c^t          & \sigma^{(k)}_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %     \right|\right)
    %     +\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \alpha_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
    % \end{align*}
    % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediately:
  \begin{align*}
    \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\sigma^{(k)}_c  - \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}  \mathbf{u}_c\right)\\
    &= \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\alpha_c\right)
  \end{align*}
